---
title: "p8105_hw2_rh3195"
author: "Ruijie He"
output: github_document
date: "2023-10-04"
---
```{r message = FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
```

## Problem 1

**Import pols-month.csv, clean data**

```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))

print(pols)
```

**Clean the data in snp.csv**

```{r clean_538_snp}
snp =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(
    year = ifelse(year >= 50, 1900 + year, 2000+year),
    month = month.name[month]) |>
  select(year, month, close)
print(snp)
```

**Tidy the unemployment.csv**

```{r tidy_538_unemployment}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
print (unemployment)
```
**Merge three dataset**

```{r}
data_538 =
  left_join(pols, snp) |>
  left_join(x =_, y = unemployment)

str(data_538)
print(data_538)
```

**Data description**

> In `pols` dataset, there are `r nrow(pols)` observations and `r ncol(pols)` variables. In `snp` dataset, there are `r nrow(snp)` observations and `r ncol(snp)` variables. And there are `r nrow(unemployment)` observations and `rncol(unemployment)` variables in `unemployment` dataset.

## Problem 2

**Read and clean Mr.Trash Wheel sheet**

```{r mr.trash wheel}
mr.trashwheel =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N586") |>
  clean_names() |>
  mutate(
    year = as.numeric(year),
    name = "mr_trash_wheel",
    homes_powered = (weight_tons * 500/30)
  )
print(mr.trashwheel)
```

**Clean professor trash wheel sheet**

```{r professor trash wheel}
professor_trash_wheel =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M108") |>
  clean_names() |>
  mutate(
    year = as.numeric(year),
    name = "professor_trash_wheel",
    homes_powered = (weight_tons * 500/30)
  )
print(professor_trash_wheel)
```

**Clean Gwynnda trash wheel sheet**

```{r Gwynnda trash wheel}
gwynnda_trash_wheel =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L157") |>
  clean_names() |>
  mutate(
    year = as.numeric(year),
    name = "gwynnda_trash_wheel",
    homes_powered = (weight_tons * 500/30)
  )
print(gwynnda_trash_wheel)
```

**Combine the three data sheet**

```{r message = FALSE}
trash_data = bind_rows(list(mr.trashwheel, professor_trash_wheel, gwynnda_trash_wheel)) |>
  select(name, everything())

print(trash_data)
```

**Data description**
> The dataset has obersations.

## Problem 3

**Import, clean, and tidy the dataset of baseline**
```{r}
baseline_data = 
  read_csv("./data/data_mci/MCI_baseline.csv", skip =1, na = c(".")) |>
  clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      0 ~ "female"
    ),
    apoe4 = case_match(
      apoe4,
      1 ~ "TRUE",
      0 ~ "FALSE")) |>
  drop_na(age_at_onset)

print(baseline_data)
```
```{r}
mean(baseline_data$current_age)
```

**Data description**
> Mean age, female amount.

**Clean and tidy amyloid dataset**

```{r}
amyloid_data =
  read_csv("./data/data_mci/mci_amyloid.csv", skip =1, na = c(".", NA)) |>
  clean_names() |>
  rename (id = study_id) |>
  pivot_longer(
    baseline:time_8,
    names_to = "time(years)",
    values_to = "amyloid_ratio"
  )
print (amyloid_data)
```

**Comment on the steps of import process**

**Join and combine baseline and amyloid data**


